---
title: "Clustering"
author: "Simon Vandekar"
date: "10/26/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(mclust)
library(Rtsne)
library(NMF)
library(parallel)
library(RColorBrewer)
library(pscl)


source('R/hierarchical_clustering.R')

knitr::knit_hooks$set(GPs=function(before, options, envir){
  if (before){
    cex=1.5
    par(mgp=c(1.7,.7,0), lwd=1.5, lend=2,
        cex.lab=0.8*cex, cex.axis=0.8*cex, cex.main=1*cex,
        mar=c(2.8,1.8,1.8,0), bty='l', oma=c(0,0,2,0))}
})
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 4, GPs=TRUE, cache=TRUE, cache.lazy = FALSE)
cols = c(brewer.pal(n=12, name='Accent'), brewer.pal(n=12, name='Dark2'), brewer.pal(n=12, name='Set3'), brewer.pal(n=8, name='Pastel2'), brewer.pal(n=9, name='Pastel1'))
transpcols = col2rgb(cols)
  transpcols = rgb(transpcols[1,], transpcols[2,], transpcols[3,], maxColorValue = 255, alpha = 127)
colscale = c('#f0f9e8', '#ccebc5', '#a8ddb5','#7bccc4','#4eb3d3', '#2b8cbe', '#08589e')
set.seed(1234)

ncores=24
```

```{r, functions, eval=TRUE}
EBupdate = function(l1s, G=3){
  
  # get level 1 parameters
  #Gzs = sapply(l1s, function(x) x$G)
  #nzs = lapply(l1s, function(x) colSums(x$z))
  # for indexing the elements from the l2 clustering
  Gzs = c(0,sapply(l1s, function(x) x$G))
  nl1s = length(l1s)
  
  
  params = do.call(rbind, lapply(l1s, function(x) t(x$parameters$mean)))
  # fit second level model
  l2 = Mclust(params, G=G)
  # probabilities from l2 model
  pzs=lapply(1:nl1s, function(l1ind) l2$z[(sum(Gzs[1:l1ind])+1):sum(Gzs[1:(l1ind+1)]),] )
  
  # higher cluster labels
  labels = lapply(pzs, function(x) apply(x, 1, which.max) )
  
  #Gz = l2$G
  
  # for testing
  # l1 = l1s[[1]]; Gz = Gzs[[1]]; pz=pzs[[1]]
  l1s = mapply(function(l1, pz, label){
    l1$parameters$mean = EBmean(mu=l1$parameters$mean,
                                sigma = l1$parameters$variance$sigma,
                                mu0 = l2$parameters$mean,
                                sigma0=l2$parameters$variance$sigma,
                                pz=pz, ns = colSums(l1$z))
    l1$l2labels = label
    l1$l2pz = pz
    ans = list(l1) }, l1=l1s, pz=pzs, label=labels)
  ans = list(l1s=l1s, l2=l2)
}

# sigma should be an array
# sigma is array of variances for l1
# sigma0 is array of variances for l2
# pz is probability of
EBmean = function(mu, sigma, mu0, sigma0, pz, ns){
  ans = do.call(cbind, lapply(1:dim(sigma)[3], function(ind){
    sig = sigma[,,ind]/ns[ind]
    mu0s = sapply(1:dim(sigma0)[3], function(ind0){
      invmat = solve(sig + sigma0[,,ind0])
      sig %*% invmat %*% mu0[,ind0] + sigma0[,,ind0] %*% invmat %*% mu[,ind]
    })
    ans = mu0s %*% pz[ind,]
  }))
}
```

## Data processing

```{r, loadData, eval=TRUE}
datafile = '/media/disk2/atlas_mxif/colon_map_20201209.rds'
atl = readRDS(datafile)
cellvars = grep('Median_Cell', names(atl), value=TRUE)
# subset out these markers that did not have
cellvars = cellvars[ - which(cellvars %in% c('Median_Cell_CD45B', 'Median_Cell_GACTIN', 'Median_Cell_PDL1', 'Median_Cell_CD45', 'Median_Cell_DAPI') )]

# remove cells not in the tumor region
atl = atl[ which(atl$Tumor>0), ]
hist(atl$Tumor)
# screen out duplicates due to FOV overlap. Eliot did this with later versions of the data, still some duplicates.
#atl = atl[ which(!duplicated(atl[,cellvars])),]
slideIDs = unique(atl$SlideID)
atl = as.data.frame(atl)

tts = read.csv('../TissueSubTypes Colon Map.csv')
# match capitalization
names(tts)[1] = 'TissueID'
atl = merge(atl, tts, all.x=TRUE)
atl$adenSSLcrc = ifelse(atl$tissueSubType %in% c('Tub', 'TV'), 'adenoma', 
                        ifelse(atl$tissueSubType %in% c('HP', 'HP/SSL', 'SSL', 'Tub/SSL'), 'SSL', 
                               ifelse(atl$tissueSubType %in% c('CRC'), 'CRC', NA )) )
# subset to non crc samples
atl = atl[ which(atl$adenSSLcrc!='CRC'),]
```


## Image normalization

```{r normalize, eval=TRUE, fig.height=14, fig.width=10}

var = cellvars[8]
# choose a subset for testing
ss = atl[sample(nrow(atl), 200000),]
# remove the NAs. Should fix this there shouldn't be any NAs
ss = na.omit(ss[,c(var, 'adenSSLcrc', 'SlideID')])
# get first slideID for each category. Model is over specified otherwise
fs = c(by(ss$SlideID, ss$adenSSLcrc, function(x) x[1]))
form = as.formula(paste(var, '~ adenSSLcrc + SlideID'))
ss$X = model.matrix(form, data=ss)
ss$X = ss$X[, grep(paste(fs, collapse='|'), colnames(ss$X), invert=TRUE, value=TRUE)]
normModel = zeroinfl(as.formula(paste(var, '~ -1 + X')), data=ss, dist = 'negbin')
# get predicted means from the model
countCoefs = coef(normModel, model='count')
ss$means = exp(ss$X %*% countCoefs)
# formula from here:
# https://ncss-wpengine.netdna-ssl.com/wp-content/themes/ncss/pdf/Procedures/NCSS/Negative_Binomial_Regression.pdf
# if negative binomial
#ss$p = ss$means/(ss$means + normModel$theta)

# plot the fit to the data in each slide
# for poisson
#by(ss, ss$SlideID, function(x){ xc= seq(min(x[,var]), max(x[,var])); hist(x[x[,var]>0,var], prob=TRUE, main=x$SlideID[1]); points(xc, dpois(xc, lambda=x$means[1]), type='l') })
# for negative binomial
layout(matrix(1:35, ncol=5, byrow=TRUE))
by(ss, ss$SlideID, function(x){ xc= seq(min(x[,var]), max(x[,var])); hist(x[x[,var]>0,var], prob=TRUE, main=x$SlideID[1]); points(xc, dnbinom(xc, mu=x$means[1], size=normModel$theta), type='l') })

# normalizes everything to the first category. Maybe not the best idea.
refmean = exp(countCoefs)



# gamma normalization in the nonzero values
nss = ss[which(ss[,var]!=0),]
form = as.formula(paste(var, '~ adenSSLcrc + SlideID'))
nss$X = model.matrix(form, data=nss)
nss$X = nss$X[, grep(paste(fs, collapse='|'), colnames(nss$X), invert=TRUE, value=TRUE)]

# doesn't fit very well
#gammaConstantVar = glm(as.formula(paste(var, '~ -1 + X')), data=nss, family=Gamma('log'))
# allows both gamma parameters to vary
library(dglm)
#gammaVarVar = dglm(as.formula(paste(var, '~ -1 + X')), as.formula(paste(var, '~ -1 + X')), data = nss, family=Gamma('log'))
#lognormalConstantVar = lm(as.formula(paste('log(', var, ')', '~ -1 + X')), data = nss)
lognormal = dglm(as.formula(paste('log(', var, ')',  '~ -1 + X')), as.formula(paste(var, '~ -1 + X')),data = nss, family = gaussian())


### FOR GAMMA
# get parameters
gammaCoefs = coef(gammaConstantVar)
nss$means = exp(nss$X %*% gammaCoefs)
disp = summary(gammaConstantVar)$dispersion
# plot the fit to the data in each slide
# for gamma
# formulas here: https://stats.stackexchange.com/questions/247624/dispersion-parameter-for-gamma-family
layout(matrix(1:35, ncol=5, byrow=TRUE))
invisible(by(nss, nss$SlideID, function(x){ xc= seq(min(x[,var]), max(x[,var])); hist(x[x[,var]>0,var], prob=TRUE, main=x$SlideID[1]); points(xc, dgamma(xc, shape=disp, rate=1/x$means[1]/disp), type='l') }))

### For Log Normal
lnCoefs = coef(lognormal)
nss$lnMeans = nss$X %*% lnCoefs
nss$sd = sqrt(exp(nss$X %*% coef(lognormal$dispersion.fit)))

layout(matrix(1:35, ncol=5, byrow=TRUE))
invisible(by(nss, nss$SlideID, function(x){ xc= seq(min(x[,var]), max(x[,var])); hist(x[x[,var]>0,var], prob=TRUE, main=x$SlideID[1]); points(xc, dnorm(log(xc), mean=x$lnMeans[1], sd=nss$sd[1])/xc, type='l') }))
```



```{r applyTransformation, eval=FALSE}
# log transform
temp = atl[,cellvars]
### COMMENT OUT TO NOT TRANSFORM THE DATA
atl[,cellvars] = log10(temp+1)# log10(sweep(temp+1, 2, colMeans(temp+1), FUN = '/') )
rm(temp)

# normalize the data by slide
cellvarsMeans = apply(atl[,cellvars], 2, function(x) mean(x[x>0]))
atl = as.data.frame(do.call(rbind, by(atl, atl$SlideID, function(data){
  zeroInds = (data[,cellvars]==0)
  # only scaling nonzero values
  datacellvarsMeans = apply(data[,cellvars], 2, function(x) mean(x[x>0]) )
  datacellvarsSDs = apply(data[,cellvars], 2, function(x) sd(x[x>0]) )
  # This makes some things negative which is not ideal for NMF.
  res = scale(data[,cellvars], center = datacellvarsMeans-cellvarsMeans, scale=FALSE)
  #res = scale(data[,cellvars], center = FALSE, scale=FALSE) # no centering or scaling
  res[zeroInds] = NA
  data[,paste0('normed_', cellvars)] = res
  data
}) ))
#atl[,paste0('normed_', cellvars)] = apply(atl[,paste0('normed_', cellvars)], 2, function(x){ans=x+min(x,na.rm=TRUE)+log10(2); ifelse(is.na(ans), 0, ans)})

# plot the normalized data


```



## Data descriptives

### Histograms

This is a histogram of the data before normalization.

```{r histBeforeTransform, fig.width=15, fig.height=18, eval=TRUE}
transpcols = col2rgb(cols)
transpcols = rgb(transpcols[1,], transpcols[2,], transpcols[3,], maxColorValue = 255, alpha = 127)
layout(matrix(1:30, nrow=6, byrow=TRUE))
#layout(matrix(1:8, nrow=2, byrow=TRUE))
invisible(sapply(cellvars,
               function(cellvar){
                 histinfo = hist(atl[,cellvar], plot=FALSE)
                 hist(atl[atl$SlideID==slideIDs[1],cellvar], breaks=histinfo$breaks, col=transpcols[1], border=NA, main=cellvar, xlab=cellvar, probability = TRUE, ylim=c(0,3)) #main=slideIDs[1]
                 sapply(2:length(slideIDs), function(ind){
                   hist(atl[ atl$SlideID==slideIDs[ind],cellvar], breaks = histinfo$breaks, border=NA, probability=TRUE, col=transpcols[ind], main=slideIDs[ind], xlab=cellvar, add=TRUE)
                 } )
                 #mtext(cellvar, outer = TRUE, cex=1.5 )
                 #dev.off()
               }))
```

These are the histograms after applying normalization
```{r, histograms, fig.width=15, fig.height=18, eval=FALSE}
layout(matrix(1:30, nrow=6, byrow=TRUE))
#layout(matrix(1:8, nrow=2, byrow=TRUE))
invisible(sapply(cellvars,
               function(cellvar){
                 histinfo = hist(atl[,paste0('normed_', cellvar)], plot=FALSE)
                 hist(atl[atl$SlideID==slideIDs[1],cellvar], breaks=histinfo$breaks, col=transpcols[1], border=NA, main=cellvar, xlab=cellvar, probability = TRUE, ylim=c(0,3)) #main=slideIDs[1]
                 sapply(2:length(slideIDs), function(ind){
                   hist(atl[ atl$SlideID==slideIDs[ind],cellvar], breaks = histinfo$breaks, border=NA, probability=TRUE, col=transpcols[ind], main=slideIDs[ind], xlab=cellvar, add=TRUE)
                 } )
                 #mtext(cellvar, outer = TRUE, cex=1.5 )
                 #dev.off()
               }))
```


### PCA
```{r, PCA, eval=FALSE}
# PCA components
scelldata = scale(atl[,cellvars])
varcomps = svd(scelldata, nu=0, nv=0)$d
varcomps = varcomps^2/sum(varcomps^2)
plot(1:length(varcomps), cumsum(varcomps), xlab='Component number', ylab='Proportion of variance', type='l')
```


### Hierarchical clustering of the markers
```{r hclust, fig.width=6, fig.height=6, eval=FALSE}
celldist <- dist(t(scelldata), method = "euclidean")
# Hierarchical clustering using Complete Linkage
hclustcell <- hclust(celldist, method = "complete" )
# Plot the obtained dendrogram
plot(hclustcell, hang = -1)
# get three clusters marker membership
members = cutree(hclustcell, k=3)
```



## Clustering

THIS NARRATIVE NEEDS TO BE UPDATED.

I've run clustering three ways:

1. Marginal clustering using naive bayes.
2. Adhoc marginal hierarchical clustering using naive Bayes. First cluster epithelium or stroma, then further categorize into cell type marginally based on known marker levels (high/low).
3. Multivariate clustering. Possibly better defined clusters, but no labels for the clusters.

Options 1 and 2 are implemented adhoc using a naive Bayes approach by averaging probabilities when multiple markers are passed. The second level of clustering is estimated only on the cells that were identified in the higher level. I used a probability threshold of .80 to select cells for building the 2nd level model. Ideally, this would instead be a weighted clustering where the weights are proportional to the cell class probabilities from the higher level.
For the first two options, zeros are removed before fitting the model and then the cells with zero values are assigned to the class with the smaller mean.
The code is not exported as a package, but could be. The function takes a `markers` data frame that contains the clustering structure.
Another potential improvement would be to use multivariate clustering when multiple markers are chosen for a given cell identifier.
This would require a more complicated decision about what latent mixture class is identified as the "expressed" class.

**Variables to change settings for the clustering are specified in this section**.

Here are the tables used to define the marginal clustering labels:

```{r clusteringParameters}
# Mclust parameters
G = 10
# Combined clustering G
Gcomb = 20
# possible models. mclust.options("emModelNames")
model='VVI' # diagonal, varying volume and shape
prior=priorControl()
Gs = seq(5, 30, by=5)
prior=priorControl()
# sample this amount for testing. Set less than 1 to not sample
nsamp = 0
nit = 5

# to assign labels
markers = data.frame(cl1=rep(c('epithelium', 'stroma'), c(5,5)),
                     cl2=c('goblet', 'tuft', 'enteroendocrine', 'stem', 'tumor', 'cytotoxicT', 'helperT', 'regulatoryT', 'B', 'Macrophages'))

### SETUP CLUSTERING PARAMETERS FOR NAIVE BAYES
markers$cl1markers = rep(list(c('Median_Cell_VIMENTIN', 'Median_Cell_PANCK', 'Median_Cell_NAKATPASE'), c('Median_Cell_VIMENTIN', 'Median_Cell_PANCK', 'Median_Cell_NAKATPASE')), c(5,5))
markers$cl1markerValue=rep(list( c(0, 1, 1), c(1,0,0)), c(5,5) )
markers$cl2markers = list('Median_Cell_MUC2', c('Median_Cell_PSTAT3', 'Median_Cell_PEGFR'), 'Median_Cell_CGA', 'Median_Cell_SOX9', 'Median_Cell_BCATENIN', c('Median_Cell_CD8', 'Median_Cell_CD3D'), c('Median_Cell_CD4', 'Median_Cell_CD3D'), c('Median_Cell_FOXP3', 'Median_Cell_CD3D'), 'Median_Cell_CD20', 'Median_Cell_CD68')
markers$cl2markerValue=list(1,c(1,1),1,1,1, c(1,1), c(1,1), c(1,1), 1,1)
markers2 = markers1 =  markers
markers2$cl1 = markers2$cl2
markers2$cl1markers = lapply(1:nrow(markers2), function(x) c(markers2$cl1markers[[x]], markers2$cl2markers[[x]]) )
markers2$cl1markerValue = lapply(1:nrow(markers2), function(x) c(markers2$cl1markerValue[[x]], markers2$cl2markerValue[[x]]) )
markers2[, grep('cl2', names(markers2))] = NULL
markersnormed = markers1
markersnormed[, c('cl1markers', 'cl2markers')] = apply(markersnormed[, c('cl1markers', 'cl2markers')], 2, function(x) lapply(x, function(y) paste0('normed_', y) ) )

### SETUP DATA FOR EACH SLIDE
slidedatas = lapply(slideIDs, function(slideID){
  slidedata = atl[ atl$SlideID==slideID , c('ID', 'Pos', 'TissueID', 'region', 'epi', 'TissueType', 'SlideID', 'x', 'y', 'tissueSubType', 'adenSSLcrc', cellvars, paste0('normed_', cellvars) )] #
  if(nsamp>0 & nsamp<nrow(slidedata)){
    slidedata = slidedata[ sample(nrow(slidedata), nsamp),]
  }
  slidedata})
names(slidedatas) = slideIDs


knitr::kable(markers1, caption = 'label1 cluster hierarchy.')
knitr::kable(markers2, caption = 'label2 cluster hierarchy.')
```


### Run Rtsne

```{r tsne, eval=FALSE}
slidedatas = mclapply(slidedatas, function(x){ x[, c('tsne_x', 'tsne_y')] = Rtsne(x[,cellvars])$Y[,1:2]; x }, mc.cores=ncores)
#slidedata = atl[atl$SlideID == slideID , cellvars]
#tsne = Rtsne(slidedata)
#s1clust = Mclust(slidedata, G=seq(5, 60, by=5), modelNames = model, prior=prior)
#plot(tsne$Y[,1], tsne$Y[,2], pch=20, col=rep(cols, 5)[s1clust$classification])
#tsneresults = paste0('tsne_results_nsamp', nsamp, '.rdata')
#save.image(tsneresults)
```


### BIC for models
Compares BIC across a range of cluster sizes specified at the beginning of this section.
```{r BIC, eval=FALSE}
BICs = mclapply(slidedatas, function(slidedata) mclustBIC(slidedata[,-1], G=Gs, modelNames = model, prior = prior), mc.cores=ncores )
BICs = lapply(BICs, function(bic){ ans = (bic-min(bic)); ans=ans/max(ans) })
plot(Gs, c(BICs[[1]]), col=cols[1], type='b', main='BIC', xlab='nclust', ylab='scaled BIC')
invisible(lapply(2:length(BICs), function(ind) points(Gs, BICs[[ind]], col=cols[ind], type='b')))
```



### Run Multivariate Gaussian mixture clustering in each slide

Fits the multivariate clustering model for each slide.

```{r mixtureModels, results='hide', eval=FALSE}
multicl = list()
for(slideID in slideIDs){
  slidedata = slidedatas[[slideID]]
  invisible(capture.output(multicl[[slideID]] <- Mclust(slidedata[,cellvars], G=G, prior=prior, modelNames = model)))
  # delete to save memory
  #mcl[[slideID]]$data = NULL
  #mcl[[slideID]]$z = NULL
}

#test = assignClasses(multicl[[1]], markers = markers2)
# prior only works with a single mean vector I think
#prior = priorControl(mean=l2$parameters$mean, modelname=model)
# EB shrinkage
# mclshrink = EBupdate(mcl, G=G)
# l2 = mclshrink[[2]]
# mclshrink = mclshrink[[1]]
# for(it in 1:nit){
#   mclshrink = EBupdate(mclshrink, G=G)
#   l2 = mclshrink[[2]]
#   mclshrink = mclshrink[[1]]
# }
# mcl = mapply(function(x, y){x$l2labels = y$l2labels; list(x)}, x=mcl, y=mclshrink)
```


### Run two level clustering in each slide

This section fits the clustering model that is coded in table `markers1`.

```{r marginalMixture, message=FALSE, results='hide', fig.width=12, fig.height=12}
  #mcl = mclapply(slidedatas, function(data){, mc.cores = ncores )
invisible(capture.output(mcl <-lapply(slidedatas, HCmulti, markers=markers1 )))
#HCmulti(slidedatas[[1]], markers1)

for(slideID in slideIDs){
    slidedatas[[slideID]][,paste0('mcl_', colnames(mcl[[slideID]][[1]]$z))] = predict(mcl[[slideID]][[1]], newdata=slidedatas[[slideID]][, colnames(mcl[[slideID]][[1]]$data)])$z
    slidedatas[[slideID]][,paste0('mcl_', colnames(mcl[[slideID]][[3]]$z))] = predict(mcl[[slideID]][[3]], newdata=slidedatas[[slideID]][, colnames(mcl[[slideID]][[3]]$data)])$z
  }
  # computes the marginal probability of each cell class
  #if(markname=='markers1') slidedatas = lapply(slidedatas, function(slide){ slide[,paste0('marg_cl2_', markers$cl2)] = slide[,paste0('cl1_', markers$cl1)] * slide[, paste0('cl2_', markers$cl2)]; slide })
```

### Mclust in full data set

```{r combinedClusteringMclust, eval=TRUE}
# Mclust parameters set above

slidedata = do.call(rbind, slidedatas)
mclfull = HCmulti(slidedata, markers=markers1[c(1,10),c('cl1', 'cl1markers', 'cl1markerValue')])
slidedata[,paste0('mclfull_', colnames(mclfull[[1]]$z))] = predict(mclfull[[1]], newdata=slidedata[, colnames(mclfull[[1]]$data)])$z

mclfull = HCmulti(slidedata, markers=markersnormed[c(1,10),c('cl1', 'cl1markers', 'cl1markerValue')])
slidedata[,paste0('mclfullnormed_', colnames(mclfull[[1]]$z))] = predict(mclfull[[1]], newdata=slidedata[, colnames(mclfull[[1]]$data)])$z

slidedatas = lapply(slideIDs, function(slideID){
    slidedata[slidedata$SlideID==slideID,]
  })
rm(slidedata)
gc()
names(slidedatas) = slideIDs
```

Running clustering by CD8 and CD4

```{r cd4cd8labels}
cd8cl = cd4cl = list()
for(slideID in slideIDs){
  slidedata = slidedatas[[slideID]]
  inds = slidedata[,'Median_Cell_CD4']>0
  invisible(capture.output(cd4cl[[slideID]] <- Mclust(slidedata[inds,'Median_Cell_CD4'], G=2, modelNames = 'E')))
  slidedatas[[slideID]][,'mcl_cd4'] = predict(cd4cl[[slideID]], newdata=slidedata[,'Median_Cell_CD4'])$z[,which.max(cd4cl[[slideID]]$parameters$mean)]
  slidedatas[[slideID]][!inds ,'mcl_cd4'] = 0
  
  slidedata = slidedatas[[slideID]]
  inds = slidedata[,'Median_Cell_CD8']>0
  invisible(capture.output(cd4cl[[slideID]] <- Mclust(slidedata[inds,'Median_Cell_CD8'], G=2, modelNames = 'E')))
  slidedatas[[slideID]][,'mcl_cd8'] = predict(cd4cl[[slideID]], newdata=slidedata[,'Median_Cell_CD8'])$z[,which.max(cd4cl[[slideID]]$parameters$mean)]
  slidedatas[[slideID]][!inds ,'mcl_cd8'] = 0
}
```


```{r writeoutLabels, eval=TRUE, cache=FALSE}
outdata = as.data.frame(do.call(rbind, slidedatas))
outdata$stromal_cells = ifelse(outdata$mcl_stroma>0.95, 1, 0)
saveRDS(outdata, file='sc_data_for_ruby.rds')
rm(outdata)
```


### Trying to see if NMF works better

```{r nmf, eval=TRUE}
slidedatas = mclapply(slidedatas, function(slidedata){
  eps = 10e-8
  x = as.matrix(slidedata[, unique(unlist(markers1$cl1markers))])
  x[ x<=0 ] = eps
  nmf.test = nmf(x=x, rank=length(unique(markers1$cl1)))
  # assigns best match marker labels
  bt = assignClasses.nmf(basis(nmf.test), coef(nmf.test), markers=markers1)
  slidedata[, paste0('nmf_', colnames(bt))] = bt
  
  # subset to the stromal class
  rpred = colnames(bt)[as.numeric(predict(nmf.test, 'rows'))]
  slidedata[,'nmf_cl1'] = rpred
  
  x = as.matrix(slidedata[ slidedata$nmf_cl1=='stroma', unique(unlist(markers1$cl2markers[ markers1$cl1=='stroma']))])
  x[ x<=0 ] = eps
  nmf.test = nmf(x=x, rank=length(unique(markers1$cl2[ markers1$cl1=='stroma'])))
  bt = assignClasses.nmf(basis(nmf.test), coef(nmf.test), markers=markers1[ markers1$cl1=='stroma',], cl = 'cl2')
  slidedata[, paste0('nmf_', colnames(bt))] = NA
  slidedata[slidedata$nmf_cl1=='stroma', paste0('nmf_', colnames(bt))] = bt
  rpred = colnames(bt)[as.numeric(predict(nmf.test, 'rows'))]
  slidedata[, 'nmf_cl2'] = NA
  slidedata[slidedata$nmf_cl1=='stroma', 'nmf_cl2'] = rpred
  slidedata
}, mc.cores = ncores )

#summary(nmf.test, class = factor(slidedatas[[1]]$epi))
#plot(slidedatas[[1]]$epi, bt[,1])
#by(bt[,1], slidedatas[[1]]$epi, quantile)

#plot(slidedatas[[1]]$epithelium, bt[,1])

```

### NMF in full data set

```{r combinedClusteringNMF, eval=TRUE}
# Mclust parameters set above

slidedata = do.call(rbind, slidedatas)
eps = 10e-8
  x = as.matrix(slidedata[, unique(unlist(markers1$cl1markers))])
  x[ x<=0 ] = eps
  nmf.test = nmf(x=x, rank=length(unique(markers1$cl1)))
  # assigns best match marker labels
  bt = assignClasses.nmf(basis(nmf.test), coef(nmf.test), markers=markers1)
  slidedata[, paste0('fullnmf_', colnames(bt))] = bt
  
  # subset to the stromal class
  rpred = colnames(bt)[as.numeric(predict(nmf.test, 'rows'))]
  slidedata[,'fullnmf_cl1'] = rpred
  
  x = as.matrix(slidedata[ slidedata$nmf_cl1=='stroma', unique(unlist(markers1$cl2markers[ markers1$cl1=='stroma']))])
  x[ x<=0 ] = eps
  nmf.test = nmf(x=x, rank=length(unique(markers1$cl2[ markers1$cl1=='stroma'])))
  bt = assignClasses.nmf(basis(nmf.test), coef(nmf.test), markers=markers1[ markers1$cl1=='stroma',], cl = 'cl2')
  slidedata[, paste0('fullnmf_', colnames(bt))] = NA
  slidedata[slidedata$nmf_cl1=='stroma', paste0('fullnmf_', colnames(bt))] = bt
  rpred = colnames(bt)[as.numeric(predict(nmf.test, 'rows'))]
  slidedata[, 'fullnmf_cl2'] = NA
  slidedata[slidedata$nmf_cl1=='stroma', 'fullnmf_cl2'] = rpred
  
  slidedatas = lapply(slideIDs, function(slideID){
    slidedata[slidedata$SlideID==slideID,]
  })
  rm(slidedata)
  gc()
  names(slidedatas) = slideIDs
```



## Plot clustering results

```{r createLabels}

for(slideID in slideIDs){
  slidedata = slidedatas[[slideID]]
  nams = paste0('mcl_', markers1$cl2[markers1$cl1=='stroma'])
  slidedata$stroma_label = as.factor(apply(slidedata[,nams], 1, function(x) nams[which.max(x)] ))
  nams2 = unique(paste0('mcl_', markers1$cl1))
  slidedata$epistr_label = as.factor(apply(slidedata[,nams2], 1, function(x) nams2[which.max(x)] ))
  nams2 = unique(paste0('mclfull_', markers1$cl1))
  slidedata$mclfull_label = as.factor(apply(slidedata[,nams2], 1, function(x) nams2[which.max(x)] ))
  nams2 = unique(paste0('mclfullnormed_', markers1$cl1))
  slidedata$mclfullnormed_label = as.factor(apply(slidedata[,nams2], 1, function(x) nams2[which.max(x)] ))
  slidedata[ slidedata$mcl_epithelium>0.5, c(nams, 'stroma_label')] = NA
  
  slidedata$mcl_cd4cat = cut(slidedata$mcl_cd4, length(colscale))
  slidedata$mcl_cd8cat = cut(slidedata$mcl_cd8, length(colscale))
  slidedatas[[slideID]] = slidedata
}
```


```{r, individualClusters, fig.width=10, fig.height=27.5, eval=FALSE}
#save.image('/media/disk2/atlas_mxif/clust_data.rdata')
#load('/media/disk2/atlas_mxif/clust_data.rdata')
for(slideID in slideIDs){
  slidedata = slidedatas[[slideID]]
  clust = mcl[[slideID]]
  l1proportions = colSums(clust$z)/sum(clust$z)
  mu = clust$parameters$mean
  rownames(mu) = cellvars
  layout(matrix(1:44, nrow=11, byrow=TRUE))
  plot(slidedata$tsne_x, slidedata$tsne_y, ylab='tsne 2', xlab='tsne 1', col=cols[slidedata$mlabel], main='Multivariate classes')
  plot(1,1, xlim=c(0,1), ylim=c(0,1), axes = FALSE, type='n', xlab='', ylab='') 
  barplot(l1proportions, col=cols, las=2, main='Class proportions')
  ## empty plots to finish the row
  plot(1,1, xlim=c(0,1), ylim=c(0,1), axes = FALSE, type='n', xlab='', ylab='') 
  
  markervals = paste0('Median_Cell_', unique(unlist(markers2$cl1markers)))
  
  invisible(sapply(c('epi', 'mcl_epithelium', 'nmf_epithelium', 'label', 'label2', markervals), function(x){ if(is.numeric(slidedata[,x]) & length(table(slidedata[,x]))>12){
    cols=brewer.pal(11, 'Spectral')
    slidedata[,x] = cut(slidedata[,x], 11)
  } else {
    cols=brewer.pal(12, 'Set3')
    slidedata[,x] = as.factor(slidedata[,x])
  }
    plot(slidedata$tsne_x, slidedata$tsne_y, ylab='tsne 2', xlab='tsne 1', main=x, col=cols[ as.numeric(slidedata[,x]) ])
    plot(1,1, xlim=c(0,1), ylim=c(0,1), axes = FALSE, type='n', xlab='', ylab='')
    legend('bottomleft', legend=levels(slidedata[,x]), fill=cols[ 1:nlevels(slidedata[,x]) ])}) )
  mtext(slideID, outer=TRUE, cex=1.5)
  # layout(matrix(1:30, nrow=6, byrow=TRUE))
  # #layout(matrix(1:8, nrow=2, byrow=TRUE))
  # trash = sapply(cellvars,
  #                function(cellvar){
  #                  histinfo = hist(atl[,cellvar], plot=FALSE)
  #                  hist(atl[atl$SlideID==slideID,cellvar], breaks=histinfo$breaks, col=transpcols[1], border=NA, main=cellvar, xlab=cellvar, probability = TRUE, ylim=c(0,3))
  #                    text(1:G, y=rep(3, G), x=mu[cellvar,])
  #                  } )
  # barplot(l1proportions, col=cols, las=2)
}
```



#### Scatterplot matrices with labels

```{r plotMatrices, fig.width=12, fig.height=12, eval=TRUE}
# invisible(lapply(slidedatas, function(slidedata, markers){
#   slideID = slidedata$SlideID[1]
#   cellvars = unique(unlist(markers$cl1markers))
#   cellvarsstroma = unique(unlist(markers$cl2markers[ markers$cl1=='stroma']))
#   cl1labels = c('epistr_label', 'mclfull_label', 'mclfullnormed_label')
#   cl2labels = c('stroma_label', 'nmf_cl2', 'fullnmf_cl2')
#   
#   for(ind in 1:3){
#     cl1lab = cl1labels[ind]
#     #cl2lab = cl2labels[ind]
#     
#     colvec = as.factor(slidedata[,cl1lab])
#     pairs(slidedata[,cellvars], panel = function(x,y) points(x,y, col=transpcols[colvec], pch=21), main=paste(slideID, cl1lab), label.pos=.75, text.panel=function(x,y, labels, cex, font){legend('bottomleft', legend=levels(colvec), fill=cols[1:nlevels(colvec)]); text(x,y, labels = labels)} )
#   }
# }, markers=markers1) )





invisible(lapply(slidedatas, function(slidedata, markers){
  slideID = slidedata$SlideID[1]
  cellvars = unique(unlist(markers$cl1markers))
  cellvarsstroma = c('Median_Cell_CD4', 'Median_Cell_CD8') #unique(unlist(markers$cl2markers[ markers$cl1=='stroma']))
  cl2labels = c('mcl_cd4cat', 'mcl_cd8cat')

  for(ind in 1:length(cl2labels) ){
    #cl1lab = cl1labels[ind]
    cl2lab = cl2labels[ind]

    colvec = as.factor(slidedata[ , cl2lab])
    pairs(slidedata[,cellvarsstroma], panel = function(x,y) points(x,y, col=colscale[colvec], pch=21), main=paste(slideID, cl2lab), label.pos=.75, text.panel=function(x,y, labels, cex, font){legend('bottomleft', legend=levels(colvec), fill=colscale[1:nlevels(colvec)]); text(x,y, labels = labels)} )
  }
}, markers=markers1) )


```


Spatial plots of the labels

```{r, fig.width=8, fig.height=8, eval=FALSE}
slidedatas = lapply(slidedatas, function(slidedata){ slidedata$epi = factor(slidedata$epi, c(1,0)); slidedata})

layout(matrix(1:4, nrow=2))
invisible(lapply(slidedatas, function(slidedata, columns){
  by(slidedata, slidedata[,c('TissueID', 'region')], function(sldt){
    slideID = sldt[1,'SlideID']
    tissueID = sldt[1,'TissueID']
    region = sldt[1,'region']
    sapply(columns, function(x){ if(is.numeric(sldt[,x]) & length(table(sldt[,x]))>12){
      cols=brewer.pal(11, 'Spectral')
      sldt[,x] = cut(sldt[,x], 11)
    } else {
      cols=brewer.pal(12, 'Set3')
      sldt[,x] = as.factor(sldt[,x])
    }
      plot(sldt$x, sldt$y, ylab='y', xlab='x', main=x, col=cols[ as.numeric(sldt[,x]) ], cex=1.5)
      legend('bottomleft', legend=levels(sldt[,x]), fill=cols[ 1:nlevels(sldt[,x]) ])}) 
    mtext(paste(slideID, tissueID, region), outer=TRUE, cex=1.5)
  })
  }, columns=c('epi', 'epistr_label', 'nmf_cl1', 'fullnmf_cl1', 'mclfull_label', 'stroma_label', 'nmf_cl2', 'fullnmf_cl2')) )
```


```{r, clusterMeans, fig.width=15, fig.height=18, eval=FALSE}
#load('/media/disk2/atlas_mxif/l2clust_data.rdata')
for(ind0 in 1:G){
  mu = l2$parameters$mean[,ind0]
  l1counts = sapply(mclshrink, function(x) sum(x$l2labels==ind0))
  #l1counts = sapply(mclshrink, function(x) sum(x$z[,x$l2labels==ind0]) )
  l1musShrink = do.call(cbind, lapply(mclshrink, function(x) x$parameters$mean[,which(x$l2labels==ind0)] ))
  l1mus = do.call(cbind, lapply(mcl, function(x) x$parameters$mean[,which(x$l2labels==ind0)] ))
  rownames(l1musShrink) = rownames(l1mus) = names(mu) = cellvars
  transpcols = col2rgb(cols)
  transpcols = rgb(transpcols[1,], transpcols[2,], transpcols[3,], maxColorValue = 255, alpha = 127)
  layout(matrix(1:30, nrow=6, byrow=TRUE))
  #layout(matrix(1:8, nrow=2, byrow=TRUE))
  invisible(sapply(cellvars,
                 function(cellvar){
                   histinfo = hist(atl[,cellvar], plot=FALSE)
                   hist(atl[atl$SlideID==slideIDs[1],cellvar], breaks=histinfo$breaks, col=transpcols[1], border=NA, main=cellvar, xlab=cellvar, probability = TRUE, ylim=c(0,3)) #main=slideIDs[1]
                     sapply(2:length(slideIDs), function(ind){
                     hist(atl[ atl$SlideID==slideIDs[ind],cellvar], breaks = histinfo$breaks, border=NA, probability=TRUE, col=transpcols[ind], main=slideIDs[ind], xlab=cellvar, add=TRUE)
                     points(y=3, x=mu[cellvar], pch=25)
                     points(y=rep(2.5, length(l1mus[cellvar,])), x=l1mus[cellvar,], pch=25, col=rep(cols, length.out=length(l1mus[cellvar,])))
                     points(y=rep(2.75, length(l1musShrink[cellvar,])), x=l1musShrink[cellvar,], pch=25, col=rep(cols, length.out=length(l1musShrink[cellvar,])))
                   } )
                   #dev.off()
                 }))
  barplot(l1counts, col=cols, las=2)
  mtext(paste('Cluster', ind0), outer = TRUE, cex=1.5 )
}
```
